---
title: "Homework 1"
author: "Hui (Sophie) Wang"
output: pdf_document
---
### 1.

#### (a) Dataset
Downloaded San Francisco Crime Classification dataset from Kaggle.
There are 148654 observations, and 13 features in the dataset.

```{r}
library(ggplot2)
salary <- read.csv("C:/Users/Sophie/Documents/Statistics/hw1/Salaries.csv")
dim(salary)
```

#### (b) Features
* Continuous features: BasePay, OvertimePay, OtherPay, Benefits, TotalPay, TotalPayBenefits, 
* Categorical features: Id, EmployeeName, JobTitle, Year, Notes, Agency, Status

Sometimes the default types R choose doesn't make sense, so I converted the types:
```{r, results='hide'}
set.seed(123)
index <- sample(1:nrow(salary), 2000)
salary2 <- salary[index,]
salary2$BasePay <- sapply(salary2$BasePay, as.numeric)
salary2$OvertimePay <- sapply(salary2$OvertimePay, as.numeric)
salary2$OtherPay <- sapply(salary2$OtherPay, as.numeric)
salary2$Benefits <- sapply(salary2$Benefits, as.numeric)
salary2$Id <- sapply(salary2$Id, factor)
salary2$Year <- sapply(salary2$Year, factor)
salary2$Notes <- sapply(salary2$Notes, as.factor)
```
```{r}
sapply(salary2, class)
```

#### (c) Summaries
Use a subset of 2000 data points to summary.
```{r}
sapply(salary2, summary)
```


#### (d) Plots
Use a subset of 2000 data points to plot.
```{r plot, fig.width=4.5, fig.height=3.8}
hist(salary2$BasePay)
hist(salary2$OvertimePay)
hist(salary2$OtherPay)
hist(salary2$Benefits)
hist(salary2$TotalPay)
hist(salary2$TotalPayBenefits)
plot(salary2$Year)
plot(salary2$JobTitle)
plot(salary2$Status)
plot(salary2$Agency)
plot(TotalPay ~ BasePay, data=salary2)
plot(TotalPay ~ Year, data=salary2)
```


### 2.

#### (a)

```{r}
set.seed(123)
index <- sample(1:nrow(diamonds), 50) # try a subset first
diamonds2 <- diamonds[index,]
summary(diamonds2$price)
hist(diamonds2$price)
```

For the normal histogram, price are divided into bins of size 2000.

* From 0 to 2000:  20 diamonds
* From 2000 to 4000: 9 diamonds
* From 4000 to 6000: 9 diamonds
* From 6000 to 8000: 4 diamonds
* From 8000 to 10000: 1 diamonds
* From 10000 to 12000: 3 diamonds
* From 12000 to 14000: 3 diamonds
* From 14000 to 16000: 1 diamonds

For the quantile histogram, price are divided into 4 quantile Since each quantile contains approximately the same number of data points, the area of each quantile is 1/4 = 0.25.

```{r}
hist(diamonds2$price, breaks=quantile(diamonds2$price))
```

* Height of 1st quantile: 0.25/(1213-537)=0.00037
* Height of 2nd quantile: 0.25/(2831-1213)=0.000155
* Height of 3rd quantile: 0.25/(5725-2831)=0.000086
* Height of 4th quantile: 0.25/(14710-5725)=0.000028

The manually calculated results match the results of R.

#### (b)
For the normal histogram, the number of diamonds with price less than 4000 is 20+9=29, and the proportion is 29/50=0.58

For the quantile histogram, the proportion of diamonds with price less than 4000 is 0.25+0.25+0.25x(4000-2831)/(5725-2831)=0.60.

The results of both histograms are about the same, if we accept some rounding errors in the computation.
